---
title: "Statistics_data"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(R.matlab)
library(BiocManager)
library(pcaMethods)
library(MetabolAnalyze)
library(MASS)
library(abind)
```

##  Load Data from csv 

```{r }

data <- as.matrix(read.csv('../data_UE1_600/data_unfold_UE1_600.csv', row.names = 1))
masked_data <- as.matrix(read.csv('../data_UE1_600/masked_data_UE1_600_04.csv', row.names = 1))

len_masked <- sum(apply(masked_data, 1, anyNA))


```

## Preprocessing 

We receive the data in a shape [x x x] and subsequently unfold it into a matrix  [`r dim(data)`]. We also separate Imaginary and Real part. Therefore, `r dim(data)[2]`/4 corresponds to the number of Multipath Components for each of the 4 sub-bands. Afterwards we split the `r dim(data)[1]` into a training and validation set, 60% and and 40% of the rows respectively. We mask the validation set with hiding 3 of the bands and liberating 1 each time. The pattern for the masking is freeing bands in this fashion: 1, 3, 2, 4 .

```{r include=FALSE}

split <- function(masked_data, len_masked){
  
  # Split Train and Validation Data in terms of the mask length 
  
  train <- masked_data[1:(nrow(masked_data) - len_masked), ]
  
  valid <- masked_data[(nrow(masked_data) - len_masked + 1):nrow(masked_data), ]
  
  list(train=train, valid=valid)
} 

train_valid <- split(masked_data, len_masked)

train <- train_valid$train
valid <- train_valid$valid

train_valid_unmasked  <- split(data, len_masked)
valid_unmask <- train_valid_unmasked$valid

#write.csv(train, 'train_data_06.csv')
#write.csv(valid_unmask, 'valid_unmask_data_04.csv')
#write.csv(valid, 'valid_masked_data_04.csv')

```


## Fitting distributions on data histograms

Here below we plot the distribution of the training and the residuals of the fitting of PCA on the training data. We overlay MLE fits (Normal and T-Student Distributions) to the histograms. 

```{r, echo=FALSE, warning = FALSE, message = FALSE}

################################################################################
################### Fitting distributions on data histograms ###################
################################################################################

fitdistnorm <- fitdistr(c(train), densfun = "normal")
fitdistt <- fitdistr(c(train), densfun = "t")

x <- seq(-.5, .5, length=100)

normfit <- dnorm(x, fitdistnorm$estimate[[1]], fitdistnorm$estimate[[2]])
hx <- dnorm(x, fitdistnorm$estimate[[1]], fitdistnorm$estimate[[2]])

tfit <- dt(x, df =  fitdistt$estimate[[3]])

yvals <- dt((c(train) - fitdistt$estimate[[1]])/fitdistt$estimate[[2]], 
            df = fitdistt$estimate[[3]])/fitdistt$estimate[[2]]

################################# Plots ########################################

hist(c(train), breaks = 500, xlim = c(-0.3, .3), freq =FALSE, 
     main = 'Histogram of the training data (UE1)')
lines(x, hx,type = 'l', # density plot ,
      col = "red")
points((c(train)), yvals, cex=.25, col='green')
legend("topright", legend=c("MLE Student-T", "MLE Normal"),
       col=c("green", "red"), lty=c(1,1), cex=0.8)
```

```{r, echo=FALSE, warning = FALSE, message = FALSE}

################################################################################
###################### Fitting distributions on the residuals ##################
################################################################################


fitdistnorm <- fitdistr(c(epsilon), densfun = "normal")
fitdistt <- fitdistr(c(epsilon), densfun = "t")

x <- seq(-.4, .4, length=100 )

normfit <- dnorm(x, fitdistnorm$estimate[[1]], fitdistnorm$estimate[[2]])
hx <- dnorm(x, fitdistnorm$estimate[[1]], fitdistnorm$estimate[[2]])

tfit <- dt(x, df =  fitdistt$estimate[[3]])
yvals <- dt((c(epsilon) - fitdistt$estimate[[1]])/fitdistt$estimate[[2]], 
            df = fitdistt$estimate[[3]])/fitdistt$estimate[[2]]

################################# Plots ########################################

hist(c(epsilon), breaks = 1500, xlim=c(-.1,.1), probability = TRUE, main="Histogram of the Error of Reconstruction of PPCA (UE1)")
lines(x, hx,type = 'l', # density plot ,
      col = "red", cex=2)
points((c(epsilon)), yvals, cex=.25, col='green')
legend("topright", legend=c("MLE Student-T", "MLE Normal"),
       col=c("green", "red"), lty=c(1,1), cex=0.8)
```

```{r, echo=FALSE, warning = FALSE, message = FALSE}
################################################################################
###################### Fitting distributions on the predicted ##################
################################################################################

fitdistnorm <- fitdistr(c(valid_pred[missing]), densfun = "normal")
fitdistt <- fitdistr(c(valid_pred[missing]), densfun = "t")

x <- seq(-1, 1, length=1000)

normfit <- dnorm(x, fitdistnorm$estimate[[1]], fitdistnorm$estimate[[2]])
hx <- dnorm(x, fitdistnorm$estimate[[1]], fitdistnorm$estimate[[2]])

tfit <- dt(x, df =  fitdistt$estimate[[3]])
yvals <- dt((c(valid_pred[missing]) - fitdistt$estimate[[1]])/fitdistt$estimate[[2]], 
            df = fitdistt$estimate[[3]])/fitdistt$estimate[[2]]

################################# Plots ########################################

hist(c(valid_pred[missing]), breaks = 400, probability = TRUE, main = 'Histogram of the Error of Reconstruction of PPCA (UE1)', xlim=c(-.3,.3))
lines(x, hx,type = 'l', # density plot ,
      col = "red", cex=2)
points((c(valid_pred[missing])), yvals, cex=.25, col='green')
legend(1, 95, legend=c("MLE Student-T", "MLE Normal"),
       col=c("green", "red"), lty=c(1,1), cex=0.8)
```

```{r, echo=FALSE, warning = FALSE, message = FALSE}
################################################################################
################ Fitting distributions on the predicted error ##################
################################################################################

fitdistnorm <- fitdistr(c(valid_pred[missing] - valid_unmask[missing]), densfun = "normal")
fitdistt <- fitdistr(c(valid_pred[missing] - valid_unmask[missing]), densfun = "t")

x <- seq(-.5, .5, length=100)

normfit <- dnorm(x, fitdistnorm$estimate[[1]], fitdistnorm$estimate[[2]])
hx <- dnorm(x, fitdistnorm$estimate[[1]], fitdistnorm$estimate[[2]])

tfit <- dt(x, df =  fitdistt$estimate[[3]])
yvals <- dt((c(valid_pred[missing]- valid_unmask[missing]) - fitdistt$estimate[[1]])/fitdistt$estimate[[2]], 
            df = fitdistt$estimate[[3]])/fitdistt$estimate[[2]]

################################# Plots ########################################

hist(c(valid_pred[missing]- valid_unmask[missing]), breaks = 1000, probability = TRUE, main = 'Histogram of the Error of prediction error of PPCA (UE1)', xlim=c(-.3, .3))
lines(x, hx,type = 'l', # density plot ,
      col = "red", cex=2)
points((c(valid_pred[missing]- valid_unmask[missing])), yvals, cex=.25, col='green')
legend("topright", legend=c("MLE Student-T", "MLE Normal"),
       col=c("green", "red"), lty=c(1,1), cex=0.8)



```