---
title: "Preprocessing + In Built"
output: html_notebook
---

```{r include=FALSE}
library(R.matlab)
library(BiocManager)
library(pcaMethods)
library(MetabolAnalyze)
library(MASS)
library(abind)
```

##  Load Data from csv 

```{r }

data <- as.matrix(read.csv('data_unfold.csv', row.names = 1))
masked_data <- as.matrix(read.csv('masked_data_04.csv', row.names = 1))

len_masked <- sum(apply(masked_data, 1, anyNA))


```

## Preprocessing 

We receive the data in a shape [x x x] and subsequently unfold it into a matrix  [`r dim(data)`]. We also separate Imaginary and Real part. Therefore, `r dim(data)[2]`/4 corresponds to the number of Multipath Components for each of the 4 sub-bands. Afterwards we split the `r dim(data)[1]` into a training and validation set, 60% and and 40% of the rows respectively. We mask the validation set with hiding 3 of the bands and liberating 1 each time. The pattern for the masking is freeing bands in this fashion: 1, 3, 2, 4 .

```{r include=FALSE}

split <- function(masked_data, len_masked){
  
  # Split Train and Validation Data in terms of the mask length 
  
  train <- masked_data[1:(nrow(masked_data) - len_masked), ]
  
  valid <- masked_data[(nrow(masked_data) - len_masked + 1):nrow(masked_data), ]
  
  list(train=train, valid=valid)
} 

train_valid <- split(masked_data, len_masked)

train <- train_valid$train
valid <- train_valid$valid

train_valid_unmasked  <- split(data, len_masked)
valid_unmask <- train_valid_unmasked$valid

#write.csv(train, 'train_data_06.csv')
#write.csv(valid_unmask, 'valid_unmask_data_04.csv')
#write.csv(valid, 'valid_masked_data_04.csv')

```

## Value Imputation through library in-built functions

Hereinafter the value imputation will be done by the in-built functions of the *pcaMethods* R library. The whole dataset without splitting is fed to the algorithm where internally imputates values.

### PPCA  

Here the function PCA is applied to use PPCA over the whole dataset. Having (NANs) inside the mentioned dataset fed to the function, NaN values are imputed through EM algorithm optimisation on for probabilistic PCA.   

```{r }


ppca_ <-  pca(masked_data, nPcs=10, method="ppca")

imputed <- completeObs(ppca_)



corrected_Abs_SE <- sqrt(mean((data[is.na(masked_data)] - 
                           imputed[is.na(masked_data)])^2) / 
                          mean(data[is.na(masked_data)]^2))

MAE <- mean(abs(data[is.na(masked_data)] - imputed[is.na(masked_data)]))
RMSE <- sqrt(mean((data[is.na(masked_data)] - imputed[is.na(masked_data)])^2))
MSE <- mean((data[is.na(masked_data)] - imputed[is.na(masked_data)])^2)

```

```{r , echo=FALSE}

table <- matrix(c(corrected_Abs_SE, MAE, RMSE, MSE), 1, 4)
rownames(table) <- c("Error Metric")
colnames(table) <- c("NRMSE", "MAE", "RMSE", "MSE")

plot(data[is.na(masked_data)], imputed[is.na(masked_data)], ylab = 'y predict', xlab = 'y true', main="Library Prediction PPCA")

knitr::kable(table)
  
```


### BPCA

Bayesian PCA, in turn, uses the algorithm detailed in the attached notes 

```{r }


bpca_ <-  pca(masked_data, nPcs=10, method="bpca")

imputed <- completeObs(bpca_)



corrected_Abs_SE <- sum((data[is.na(masked_data)] - imputed[is.na(masked_data)])^2) / var(data[is.na(masked_data)])

MAE <- mean(abs(data[is.na(masked_data)] - imputed[is.na(masked_data)]))
RMSE <- sqrt(mean((data[is.na(masked_data)] - imputed[is.na(masked_data)])^2))
MSE <- mean((data[is.na(masked_data)] - imputed[is.na(masked_data)])^2)


```

```{r , echo=FALSE}

table <- matrix(c(corrected_Abs_SE, MAE, RMSE, MSE), 1, 4)
rownames(table) <- c("Error Metric")
colnames(table) <- c("NRMSE", "MAE", "RMSE", "MSE")

knitr::kable(table)
  

```

